refnx - Neutron and X-ray reflectometry analysis in Python

### 2018-07-21 paper/manuscript.tex
PAPER: generalised curvefitting
PAPER: how long it takes

The powerful curve-fitting engine offers the choice of using a normal least-squares fitting approach (with a variety of global and gradient-based optimizers) or a Bayesian approach using Markov Chain Monte Carlo to investigate the posterior distribution of the model parameters. The Bayesian approach is useful in examining parameter covariances, model selection, and variability in the resulting scattering length density profiles.

@@

The \Objective\ statistics are used directly by the \textbf{CurveFitter} class to perform least-square fitting or Bayesian MCMC of the system. \refnx\ uses the \SciPy\ package for least-squares analysis (DifferentialEvolution, Levenberg Marquardt, LBFGSB - Limited Broyden–Fletcher–Goldfarb–Shanno with bounds) and the  \emcee\ package \cite{emcee} for affine invariant MCMC ensemble sampling of the posterior distribution of the parameters. The likelihoods used in MCMC and least-squares assume that the measurement errors are normally distributed, Equation~\ref{eqn:2}. However, other types of measurement errors (e.g. Poissonian) could be implemented by a subclass of \Objective\  overriding the log-likelihood method.
Parallelisation of the sampling is done automatically, and can use MPI on a cluster for greater parallelisation. Visualisation of the samples produced by MCMC sampling is done using the \corner\ package \cite{corner} for scatter plot matrices. On the conclusion of a fit/sampling run each \Parameter\ value is updated and given a standard error. For the sampling these represent the median and half the [15.87, 84.13] percentile range respectively; the latter approximating the standard deviation for a normally distributed statistic.

@@

\begin{figure}
  \includegraphics[width=\linewidth]{./supporting_information/corner}
  \caption{Corner plot for the varying parameters of DMPC bilayers measured at three contrasts.}
  \label{fig:corner}
\end{figure}

>>>>

The powerful curve-fitting engine offers the choice of using a normal least-squares fitting approach (with a variety of global and gradient-based optimizers) or a Bayesian approach using Markov Chain Monte Carlo to investigate the posterior distribution of the model parameters. The Bayesian approach is useful in examining parameter covariances, model selection, and variability in the resulting scattering length density profiles. Whilst the main intention of the program is to analyse reflectometry data it is equally adept for generalised curve-fitting purposes.

@@

The \Objective\ statistics are used directly by the \textbf{CurveFitter} class to perform least-square fitting or Bayesian MCMC of the system. \refnx\ uses the \SciPy\ package for least-squares analysis (DifferentialEvolution, Levenberg Marquardt, LBFGSB - Limited Broyden-€“Fletcher-€“Goldfarb-€“Shanno with bounds) and the  \emcee\ package \cite{emcee} for affine invariant MCMC ensemble sampling of the posterior distribution of the parameters. The likelihoods used in MCMC and least-squares assume that the measurement errors are normally distributed, Equation~\ref{eqn:2}. However, other types of measurement errors (e.g. Poissonian) could be implemented by a subclass of \Objective\  overriding the log-likelihood method.
Parallelisation of the sampling is done automatically, and can use MPI on a cluster for greater parallelisation. Visualisation of the samples produced by MCMC sampling is done using the \corner\ package \cite{corner} for scatter plot matrices. On the conclusion of a fit/sampling run each \Parameter\ value is updated and given a standard error. For the sampling these represent the median and half the [15.87, 84.13] percentile range respectively; the latter approximating the standard deviation for a normally distributed statistic.

@@

\begin{figure}
  \includegraphics[width=\linewidth]{./supporting_information/corner}
  \caption{Corner plot for the varying parameters of DMPC bilayers measured at three contrasts. The sampling took $\sim$12 minutes for 30 saved steps, corresponding to 6000 samples, with the steps being thinned by a factor of 100.}
  \label{fig:corner}
\end{figure}
